\documentclass[a4paper]{report}

% Number of "levels" the label can reference. Default is 2.
\setcounter{secnumdepth}{3}

\input{settings}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Header and Footer Setup %
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\lhead{Alberte Thegler}
% \chead{}
\rhead{\leftmark}

\renewcommand{\headrulewidth}{0.4pt} % thickness of line at header
\renewcommand{\footrulewidth}{0.4pt} % thickness of line at footer
\setlength{\belowcaptionskip}{0pt} % space below captions


%%%%%%%%%%%%%%%%%%%%
% Front-page Setup %
%%%%%%%%%%%%%%%%%%%%
%
\assignment{Master's Thesis}
\author{Alberte Thegler - alberte@thegler.dk}
\title{Towards Automatic Program Specification \\ From SME Models}
\subtitle{Department of Computer Science}
\date{Handed in: November 21, 2018}
\advisor{Supervisors: Brian Vinter and Kenneth Skovhede}


\begin{document}

\maketitle

\pagenumbering{roman}


\begin{abstract}
\begin{doublespace}
This thesis introduce a method to simplify hardware modeling and verification
thereof, in order for software programmers to, more easily, meet the demands of
the growing embedded device industry. The system TAPS is presented, which is a
system for transpiling from the new SME Implementation Language (SMEIL) into the
machine-readable CSP language (\cspm{}).
The translated \cspm{} code is augmented with assertion statements, derived from values observed by simulating SME models. Utilising the formal verification
properties of the FDR4 refinement tool, the assertion statements of the
translated \cspm{} code can be formally verified.
An extension to the initial version of TAPS is also introduced in the thesis. The extension provides the possibility of modeling a globally synchronous clock in
\cspm{}, providing the possibility of verifying the internal state of the
\cspm{} network separately for each clock cycle.
A small example consisting of a
seven segment display clock network is presented, to introduce how to verify the
widths of channels in the network. A small cyclic network is presented, to
show the possibilities of verifying separate clock cycles, using the extended
version of TAPS.
\end{doublespace}
\end{abstract}

\chapter*{Acknowledgements}
I would like to thank my supervisors Brian Vinter and Kenneth Skovhede for their support, patience, and guidance throughout this project, and for encouraging me to write a paper based on this work.
I'd like to give special thanks to Mads Ohm Larsen for his immense help in writing the paper and for always taking the time to discuss my ideas.
Another special thanks goes to Truls Asheim for his kind support in introducing me to SMEIL, proofreading my writings, and for patiently answer all of my questions.
I would also like to thank the employees in the eScience group at the Niels Bohr Institute, for their encouragement and for providing a welcomming environment.
Last but not least, I would like to thank Lars and Asbj\o rn Thegler for taking their time to proofread my thesis and for their help towards making my writings clear and concise.

\newpage
\tableofcontents

\newpage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%
% Include Content %
%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\input{chapters/introduction}

\chapter{Related Work}
\label{chap:related_work}
\input{chapters/related_work}

\chapter{Approach}
\input{chapters/approach}

\chapter{Background Technologies}
\label{chap:background}
\input{chapters/background}

\chapter{Analysis}
\label{chap:analysis}
\input{chapters/analysis}

\chapter{Designing TAPS}
\label{chap:design}
\input{chapters/design}

\chapter{Implementation}
\label{chap:implementation}
\input{chapters/implementation}

\chapter{Extending TAPS for Clocked Systems}
\label{chap:clock}
\input{chapters/clock_cycle_problem}

\chapter{Experiments and Results}
\label{chap:exp}
\input{chapters/experiments_and_results}

\chapter{Discussion}
\label{chap:discussion}
\input{chapters/discussion}

\chapter{Future Work}
\label{chap:future}
\input{chapters/future_work}

\chapter{Conclusion}
\input{chapters/conclusion}

% \chapter*{Acknowledgements}
% % Thanks to Uwe Zimmermann who made the seven segment example in TikZ on \url{http://www.texample.net/tikz/examples/segment-display/}.


% \newpage
% citation for citations\cite{Bengtsson1995}


%%%%%%%%%%%%%%%%
% Bibliography %
%%%%%%%%%%%%%%%%
\newpage
\bibliographystyle{abbrv}
\bibliography{library}

%%%%%%%%%%%%%%%%%%%%
% Include Appendix %
%%%%%%%%%%%%%%%%%%%%
\appendix
\input{chapters/appendix}

\end{document}
